### 2024-03-31 output/paper/paper.qmd
Uploading quarto file

---
title: "Analyzing the Relationship between Female Fertility Rate and Female Literacy Rate in South Asia"
author: "Jagpreet Singh"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
date-format: long
bibliography: references.bib
execute: 
  echo: false
format: pdf
fig-pos: "H"
abstract: TBD
---

```{r}
knitr::opts_chunk$set(echo = FALSE, results ="hide", message = FALSE, warning = FALSE)
```

# Introduction

# Data {#sec-data}

## Data Source {#sec-variables-of-interest}

## Variables of Interest

# Model {#sec-model}

## Model Justification

# Results {#sec-results}

# Discussion {#sec-discussion}